Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
delete duplicate chunk, minor reference fixes

\subsection{Main Results}
These are our results.
\begin{thm}
\label{thm:g_0_generators_relations}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

@@

\section{High genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Sections ~\ref{sec:g_1} and ~\ref{sec:g_0_effective} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

@@

\begin{enumerate}
	\item $\halfcan' = 0$
	\item $\halfcan' = P - Q + \frac{1}{3}R_1 + \frac{1}{3}R_3.$
\end{enumerate}
\todo{make table}

@@

Then, there exist constants $\alpha,\beta,\gamma \in k$ so that 
$R_{\halfcan'} \cong k[u,x,y_3,y_4]/(xy_3- \alpha uy_4,y_4^2 - \beta x^2 u - \gamma y^2u).$ The proof of this is fairly algorithmic, and analogous to that completed in Example ~\ref{ex:base-377}.

@@

\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fixed points, and $R_1,\ldots, R_n$ be points on $\sx$. In this section, will compute the canonical ring $R_{\halfcan'}$ in the cases that $\sigma = (1;-;0),\sigma = (1;e;0)$. That is, we will consider the cases.
\begin{enumerate}
	\item $\halfcan' = P-Q$
	\item $\halfcan' = P - Q + \frac{1}{3}R$
	\item $\halfcan' = P - Q + \frac{1}{5}R$
	\item $\halfcan' = P - Q + \frac{e-1}{2e}R, e > 5, e \equiv 1 \bmod 2$ 
\end{enumerate}
\todo{make table}

\begin{example}
\label{eg:exceptional-1-0}
In the case $L=P - Q.$ Then, we have $R_{(2k+1)L} \cong 0$ while $R_{2kL} \cong k$, with a single generator in degree 2. Therefore, $R_L \cong k[y]$, where $y$ is a generator in degree 2. 
\end{example}

\begin{example}
\label{eg:exceptional-1-3}
Let $L = P - Q + \frac{1}{3}R$. Since $P \neq Q$, we see $h^0(L) = 0,$ and we know $h^0(2L) = h^0(K) = 1$. For $n > 2$, we have $\deg nL > 0$ and so by Riemann--Roch, we have $h^0(nL) = \deg \lfloor nL/3 \rfloor.$ Next, 
We find that $R_L \cong k[u,x,z]/(z^2 - ux^4),$ where $\deg u = 2, \deg x = 3, \deg z = 7$ so that $\di u|_L = -2P + 2Q, \di x|_L = -3P + 3Q - R, \di z|_L = -7P + 7Q - 2R.$\todo{Peter: Why don't we have something in degree 4?}
\end{example}

\begin{example}
\label{eg:exception-1-5}
Let $L = P - Q + \frac{2}{5}R_1.$ Here, we 
\todo{say what the variables are}
$R_L \cong k[u,y_3, y_5]/(y_3^4 - \alpha uy_5^2),$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3,\ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\todo{Check admissibility}
\end{example}

\begin{example}
\label{eg:exception-1-e}
Let $L = P - Q + \frac{e-1}{2e}R_1.$ Here, we claim $R_L \cong k[u,y_3, y_5, \ldots, y_e]/I_e,$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3,\ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\end{example}

@@

\section{Genus Zero Effective}
\label{sec:g_0_effective}
Now we consider the genus zero case.
\todo{write intro}

@@

\section{Genus Zero Non-effective}
\label{sec:g_0_non_effective}

@@

In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations generated in degree at most $2e$, so long as $\sigma$
does not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_noneff_induction}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

@@

The following table demonstrates that $R(\sx', \Delta, L')$ is
generated as a $k$-algebra by elements of degree at most $e$
with relations in degree at most $2e$ for each case. Also note
that in these cases, $e_i = e + 2 > \deg z$ for all $i \in J$
and any generator $z$ of $R'$.

\begin{longtable}
	{| c || c | c | c |}
	\hline
	Case & Generator Degrees & Minimal Relation Degrees & $e$\\
	\hline
	\hline

@@

\begin{longtable}
	{| c || c | c | c |}
	\hline
	Signature & Generator Degrees & Minimal Relation Degrees & $e$ \\
	\hline
	\hline

@@

\begin{center}
\textbf{Table (IV)}: Genus 0 exceptional cases
\end{center}

@@

\begin{thm}
\label{thm:g_0_generators_relations_rep}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

>>>>

\subsection{Main Results}
\begin{enumerate}
	\item Petri's Theorem
	\item Modular Forms
	\item Stacky Relations
	\item Statement of Results
	\item Outline of paper
\end{enumerate}

These are our results.
\begin{thm}
\label{thm:g_0_generators_relations}
Let $(\sx, \Delta, \halfcan)$ log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d = 0}^\infty H^0(\sx, \sco(L)^{\otimes d})
\end{align*}

\noindent
is generated as a $k$-algebra by elements of degree at most $e = 
\max(5,e_1,\ldots, e_r).$ If the genus $g \geq 2,$ then $R(\sx, \Delta, \halfcan)$ has relations in degree at most $\max(11,e_1,\ldots, e_r).$ If the genus $g = 0$ or $g = 1,$ then $R(\sx, \Delta, \halfcan)$ has relations in degree at most $\max(10,e_1,\ldots, e_r),$ so
long as $\sigma$ does not lie in a finite list of exceptional cases, as listed in Table ~\ref{table:g-1-exceptional} for signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional} for signatures with $g = 0$.
\end{thm}

Note that in the course of the proof of Theorem ~\ref{thm:g_0_generators_relations} we will not merely obtain bounds on the degrees of the generators and relations. In fact, if the genus $g = 0$ or $g = 1$, the proof yields an explicit system of generators and the initial ideal of relations. While in the genus $g \geq 2$ case, the proof does not always yield an explicit system of generators and relations, if the halfcanonical divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',0,X)$ a log spin curve, and a system of generators and relations for $\halfcan'$ is known, then our proof provides an inductive procedure for determining the generators and initial ideal of relations for $\halfcan$. In fact, such a system of generators and relations are known for $\halfcan'$ in many cases, as detailed in Neves ~\cite{neves:halfcan}[Section III.4].

@@

\section{High genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Section ~\ref{sec:g_1} and Subsection ~\ref{ssec:g_0_effective} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

@@

\begin{enumerate}
	\item $\halfcan' = 0$
	\item $\halfcan' = P - Q + \frac{1}{3}R_1 + \frac{1}{3}R_3.$
\end{enumerate}

@@

Then, there exist constants $\alpha,\beta,\gamma \in k$ so that 
$R_{\halfcan'} \cong k[u,x,y_3,y_4]/(xy_3- \alpha uy_4,y_4^2 - \beta x^2 u - \gamma y^2u).$ The proof of this is fairly algorithmic: We may first use Riemann--Roch to compute the dimensions of $(R_{\halfcan'})_n$ over $k$, then verify that these generators and relations produce the correct number of independent functions via an analysis of zeros and poles. The details are omitted as it is analogous to Example ~\ref{ex:base-377}.

@@

\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fixed points, and $R$ be a point on $\sx$.

\todo{cite genus 1 final theorem here, to explain what these cases are exceptional to}
\begin{longtable}	{| c || c | c | c |}
	\hline
	L' & Generator Degrees & Degrees of Minimal Relations \\
	\hline
	\hline
	$(1;-; 0)$ & $\{2\}$ & $\emptyset$ \\	\hline

	$(1; 3; 0)$ & $\{2,3,7\}$ & $\{14\}$ \\	\hline

	$(1; 5; 0)$ & $\{2,3,5\}$ & $\{12\}$ \\	\hline

%	$(1; e; 0)$ & $\{2, 3, 5, 7, 9, \ldots, e\}$ & $\{12\} \cup \{9,10,\ldots, e+2, e+3, e+5,e+7,\ldots, 2e-4\}$ \\	\hline
\end{longtable}

\begin{center}
\label{table:g-1-exceptional}
\textbf{Table (IV)}: Genus 1 exceptional cases, where $L' = P-Q + \frac{e-1}{2e}R$, with $P \neq Q$.
\end{center}
\iffalse
\begin{rem}
\label{rem:g-1-exceptional-e}
Note that for $e > 5$ generators and relations for $L = P -Q + \frac{e-1}{2e}R$, which has signature $(1;e;0)$, can be computed inductively $L' = P-Q + \frac{(e-2)-1}{2(e-2)}R$ using 
Theorem ~\ref{thm:ram_order_ind}.
\end{rem}
\fi

We now elaborate on the computation of the results in 
Table ~\ref{table:g-1-exceptional}. In the following example we will explain the case $\sigma = (1;5;0)$ as the other cases are even easier.

\iffalse
\begin{example}
\label{eg:exceptional-1-0}
In the case $L=P - Q.$ Then, we have $R_{(2k+1)L} \cong 0$ while $R_{2kL} \cong k$, with a single generator in degree 2. Therefore, $R_L \cong k[y]$, where $y$ is a generator in degree 2. 
\end{example}

\begin{example}
\label{eg:exceptional-1-3}
Let $L = P - Q + \frac{1}{3}R$. Since $P \neq Q$, we see $h^0(L) = 0,$ and we know $h^0(2L) = h^0(K) = 1$. For $n > 2$, we have $\deg nL > 0$ and so by Riemann--Roch, we have $h^0(nL) = \deg \lfloor nL/3 \rfloor.$ Next, 
We find that $R_L \cong k[u,x,z]/(z^2 - ux^4),$ where $\deg u = 2, \deg x = 3, \deg z = 7$ so that $\di u|_L = -2P + 2Q, \di x|_L = -3P + 3Q - R, \di z|_L = -7P + 7Q - 2R.$\todo{This is unfinished}
\end{example}

\begin{example}
\label{eg:exception-1-e}
Let $L = P - Q + \frac{e-1}{2e}R_1.$ Here, we claim $R_L \cong k[u,y_3, y_5, \ldots, y_e]/I_e,$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3,\ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\end{example}
\fi

\begin{example}
\label{eg:exception-1-5}
Let $L = P - Q + \frac{2}{5}R_1$ be a half canonical divisor on a stacky curve $\sx$. Let $u \in (R_L)_1$ be any nonzero element. We obtain $\di u|_R = 0,$ since $2P -2Q \sim 0$ and by Riemann--Roch $\dim_k (R_L)_1 = 1$ let $y_3 \in (R_L)_3$ be any nonzero element. We obtain $\di y_3|_R = - R,$ by Riemann Roch, since if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0(\sx,3P-3Q) \cong H^0(\sx,P-Q) \cong 0$ as $P \neq Q$. Finally, let $y_5 \in (R_L)_5$ be an element with $y_5|_R = -R$. Then, we claim there is some $\alpha \in k$ so that
$R_L \cong k[u,y_3, y_5]/(y_3^4 - \alpha uy_5^2).$
The proof of this can be gotten by first using Riemann--Roch and counting the pole order at $R$ considerations to check the above relation exists, and that the generators and relation determine a ring whose $k$th graded piece has the same dimension of $R_L$. It is analogous to Example ~\ref{ex:base-377}, and is omittied.
\end{example}

@@

\section{Genus Zero}
\label{sec:g_0}
Now we consider the genus zero case.
\todo{write intro}

\ssec{Genus Zero Effective}
\label{ssec:g_0_effective}

@@

\ssec{Genus Zero Non-effective}
\label{ssec:g_0_non_effective}

@@

In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations generated in degree at most $2e$, so long as $\sigma$
does not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_noneff_induction}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} \todo{fix} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

@@

The following table demonstrates that $R(\sx', \Delta, L')$ is
generated as a $k$-algebra by elements of degree at most $e$
with relations in degree at most $2e$ for each case. Also note
that in these cases, $e_i = e + 2 > \deg z$ for all $i \in J$
and any generator $z$ of $R'$.
\begin{longtable}
	{| c || c | c | c |}
	\hline
	Case & Generator Degrees & Degrees of Minimal Relations & $e$\\
	\hline
	\hline

@@

\begin{longtable}
	{| c || c | c | c |}
	\hline
	Signature & Generator Degrees & Degrees of Minimal Relations & $e$ \\
	\hline
	\hline

@@

\begin{center}
\label{table:g-0-exceptional}
\textbf{Table (IV)}: Genus 0 exceptional cases
\end{center}

@@

\todo{Aaron: Use thmrestate instead of rewriting the Theorem}
\begin{thm}
\label{thm:g_0_generators_relations_rep}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring
